### 2017-12-02 thesis/contents/reviews/dmri.tex
checkpoint on edit

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=3.5in]{diffusion.pdf}
\end{tabular}
\caption{Isotropic diffusion: diffusion of molecules have the same probable displacement in all directions. Anisotropic diffusion displacement of modules are constrained to its diffusion medium.}
\label{fig:diffusion}
\end{center}
\end{figure}

@@

To properly characterize anisotropic diffusion, there must be more than 1 gradient directions available. In fact it is possible to describe voxel-wise diffusion using a Gaussian diffusion model, where the diffusion anisotropy can be visualized as an ellipsoidal tensor \cite{Jones2002}. This type of imaging methodology is termed diffusion tensor imaging (DTI). At least 6 non-collinear gradient directions must be captured in order to compute the DT. In practice up to 32 gradients directions are used clinically.

The diffusion tensor is a $3 \times 3$ symmetric matrix that describes diffusion displacement in the three major axis of diffusion:

@@

The eigen values are measured in the same unit as ADC, which is in $mm^2/s$. Their corresponding eigen vector is denoted as $\hat{\lambda_i}$ They are ordered from the greatest to the least diffusivity magnitude. $\hat{\lambda_1}$ points to the orientation of the greatest diffusion direction, and therefore $\lambda_1$ is also referred to as "longitudinal" or "axial" diffusivity. $\hat{\lambda_2}$ and $\hat{\lambda_3}$ are the lesser orthogonal diffusion directions, and their mean magnitude give rise to the cross-sectional, or "radial" diffusivity.

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=2.5in]{tensor.pdf}
\end{tabular}
\caption{Illustration of the eigen vectors of an ellipsoid tensor.}
\label{fig:tensor}
\end{center}
\end{figure}

@@

FA has been found to correlate with a wide spectrum of neuromorphological changes, including but not limited to multiple sclerosis \cite{Cercignani2001,Ciccarelli2003d,Miron2012}, vestibular schwannomas \cite{Chen2011b,Taoka2006,Wai2009}, tumors \cite{Mori2002,Schonberg2006,Nimsky2005}, schizophrenia \cite{Rametti2009,Fitzsimmons2009}, traumatic brain injury \cite{Bendlin2008,Kinnunen2011c,Hellyer2012}.

RD is found to positively correlate with degree of demyelination in multiple sclerosis \cite{Song2005,Song2002,Klawiter2011,Janve2013}, and well as stages of white-matter development and aging \cite{Counsell2006,Sala2010}, including alzheimer's disease \cite{Acosta-Cabronero2012}.

AD is correlated positively with axonal integrity \cite{Song2003}, and is found to be a correlate of axonal injury \cite{Budde2009} such as those found in traumatic brain injury \cite{Kinnunen2011c} and Gulf War illness \cite{Rayhan2013},

MD is correlated with tissue inflammation and edema \cite{Senda2012g}, and is found to correlate with diffusivity changes as a result of gliomas \cite{Stadlbauer2007}, multiple sclerosis \cite{Cercignani2001,Senda2012g,DeGroot2013}, amyotrophic lateral sclerosis \cite{Sharma2012}, and aging \cite{Kantarci2011}.

@@

In order to characterize the additional diffusion components in voxels with multiple fibres, changes in DWI acquisition strategies is necessary. A natural extension is to acquire additional diffusion directions. This type of acquisition strategies, which enable higher order diffusion modelling algorithms, are generally referred to as high angular resolution diffusion imaging (HARDI). The most wide spread strategy is to equally distribute gradient directions as vertices of a triangle-tessellated geodesic sphere \cite{Frank2001}. Alternatively, it's also possible to equally spread gradient directions in a rectilinear grid, which forms the basis of diffusion spectral imaging (DSI) \cite{Wedeen2005}. Additional acquisition strategies such as distributing gradient directions across multiple concentric spherical shells is also actively explored \cite{Jeurissen2014a}.

@@

The determination of gradient sampling pattern and B-value directly affect acquisition time and image quality. More complex gradient sampling and higher B-values can improve tractography quality, however the acquisition time can become prohibitively long in a clinical setting. Other factors that must be considered include the nature of patient pathology, motion artifacts, and the inclusion of other imaging modalities during clinical scan sessions.

@@

An end-to-end spherical deconvolution diffusion model was first proposed by Tournier et al \cite{Tournier2004}. 
Consider a fiber population with distinct fiber components. The measured diffusion signal $S(\theta,\phi)$ can be written as:

\begin{equation}
S(\theta, \phi) = \sum_{i}^{} f_i \hat{A}_i R(\theta)
\end{equation}
where $\theta$ is the elevation angle,  $\phi$ the azumuth angle in spherical coordinates; $f_i$ is the volume fraction of the $i$-th fiber component, $\hat{A}$ is the rotation operator that orients fiber, and $R(\theta)$ is a axially symmetric response function.

@@

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=100mm]{harmonics.pdf}
\end{tabular}
\caption{Visualization of the spherical harmonic functions up to l=3.}
\label{fig:harmonics}
\end{center}
\end{figure}

Spherical decovolution is considered to be the action of a set of rotations on a function over a sphere. Given $F^l $ as a vector of $l$th order spherical harmonic decomposition of  $F(\theta,\phi)$, with vector length $(2l + 1)$, and $R^l$ as the $l$th order rotational harmonic decomposite of $R(\theta)$, with vector length $(2l+1)^2$. The spherical harmonic representation of $S(\theta,\phi)$ is then:
\begin{equation}
S^l = R^l F^l
\end{equation}
To reduce the number of parameters to be computed, we can make the further assumption that the diffusion process is symmetric, and therefore the odd $l$ harmonic orders can be ignored, with only the even harmonics considered. 
The response function $R(\theta)$ can be directly estimated from the dataset by supplying DWI region with high anisotropy from a single fiber bundle, such as the body of corpus callosum.

The maximum haronic order $l_max$ that can be estimated is dependent on obtained number of directions. For example, $l_max$ = 8 requires 45 directions, $l_max$ = 10 requires 66 directions, and so on. Larger $l_max$ are more sensitive to noise, which is a weakness for the first spherical deconvolution algorithm \cite{Tournier2004}.

\subsubsection{Constrained Spherical Deconvolution}
Further advances to the estimation method of the spherical deconvolution model made it more robust to noise \cite{Tournier2007b}. The noise of SD estimation are represented as spurious negative lobes in FOD, which are physically impossible, and do not represent any real phenomenon. To mitigate this, a non-negativity constraint is added. This method is referred to as \textit{constrained spherical deconvolution} (CSD).

The CSD constraint is applied iteratively using Tikhonov regularization \cite{Hansen1998}. The procedure involves obtaining an initial FOD estimate, and a set of directional peaks are then identified. The Tikhonov regularization then constraints the negative peaks to zero, and a new set of FOD and peaks are obtained. This is applied iteratively until convergence, and the final constrained FOD is then supplied.

The Tikhonov regularization minimizes the weighted sum two terms:
\begin{equation}
\|{Af - b}^2\|^2 + \lambda^2 \| l(f - f')\|^2
\end{equation}
Where the first term is a lear least-square fit of solution $f$ to to data $b$ with problem matrix $A$. 
In terms of CSD, $f$ corresponds to the SH coefficients of the FOD, $b$ is the measured signal and $A$ is:
\begin{equation}
A = QR
\end{equation}
Where $Q$ is the SH coefficient to DW intensity mapping, and $R$ is the spherical deconvolution. 
The regularization term is the constrained matrix $L$ multiplying the norm of $f$ relative to an initial estimate $f'$. The parameter $\lambda$ controls the regularization amount, with $\lambda=0$ having no regularization, to $\lambda\rightarrow \infty $ making the solution tend to $f'$.

@@

CSD is robust to noise at relatively low B-values, and is capable of good quality FOD estimates at $B=1000 s/mm^2$. This means that existing DTI protocols can be extended for high order HARDI tractography by simply increasing the number of directions. With CSD, a 60 direction DWI acquisition can be used to confidently derive FOD with $l_{max}=8 $, and up to $l_{max}=12$ with super resolution enabled \cite{Tournier2007b} (\ref{fig:csd}).

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{csd-angles}
\end{tabular}
\caption{ Comparison of angle discrimination ability and CSD FOD. B-value=3000 \protect $s/mm^2$. Taken from \protect \cite{Tournier2007b} } 
\label{fig:csd}
\end{center}
\end{figure}

\subsection{Tractography}
Tractography is the construction, or propagation of streamline from an underlying field of orientation information, in order to visualize continuous structures within a volume.  The orientation field is generally provided as a 3D matrix of directional information obtained from DTI, CSD, or other methods. For example, in DTI the directions is the eigen vector with the greatest magnitude $\hat{\lambda_1}$; for CSD, the directional vectors would be peaks from the FOD.

Applying pathway tracking to DWI imaging was first explored by Mori \cite{Mori2002b} and Basser\cite{Basser2000}. The family of methods that propagates by deterministic algorithms based on local orientation data is collectively called deterministic tractography. The first tracking algorithms were applied to DTI using step-wise polyline propagators that followed local DT tensors \cite{Conturo1999}. Further refinement to the line propagator to include neighbouring DT tensors, and interpolation of the local DT directions in order to obtain a smooth curve \cite{Mori2002b}. Basser introduced Runge-Kutta integration to allow more robust path tracing when faced with noise in the local directions \cite{Basser2000}. Another similar path-integral method was also proposed by Tuch \cite{Tuch2000d}. The integrated tractography approaches are now standard practice for deterministic tractography \cite{Tournier2004,Pieper2004}.

Deterministic tractography lends itself naturally to anatomical interpretation and further geometric analysis. It was traditionally paired with DTI, and so was thought to be limited in the case of crossing fibers. Other methods attempted to circumvent the limitation with probabilistic approaches that include white matter tissue priors in the fiber propagation model. The most popular of the probabilistic approach is the FSL software's probabilistic tractography module (FDT) \cite{Behrens2007}, which also includes a model-based ball-and-stick diffusion model that can resolve 2-fiber crossings. The FDT software produces visitation images of its probabilistic streams that collates the probability of tract propagation at each voxel.  The visitation image however is hard to interpret and localize, and therefore does not rival deterministic methods in surgical visualization.

@@

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{tractography1}
\end{tabular}
\caption{Example of cranial nerve reconstruction with deterministic tractography.} 
\label{fig:tract1}
\end{center}
\end{figure}

>>>>

\begin{figure}[ht]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=3.5in]{diffusion.pdf}
\end{tabular}
\caption{Isotropic diffusion: diffusion of molecules have the same probable displacement in all directions. Anisotropic diffusion displacement of modules are constrained to its diffusion medium.}
\label{fig:diffusion}
\end{center}
\end{figure}

@@

To adequately characterize anisotropic diffusion, there must be more than one gradient directions available. In fact, it is possible to describe voxel-wise diffusion using a Gaussian diffusion model, where the diffusion anisotropy can be visualized as an ellipsoidal tensor \cite{Jones2002}. This type of imaging methodology is termed diffusion tensor imaging (DTI). At least six non-collinear gradient directions must be captured to compute the DT. In practice, up to 32 gradients directions are used clinically.

The diffusion tensor is a $3 \times 3$ symmetric matrix that describes diffusion displacement in the three major axes of diffusion:

@@

The eigen values are measured in the same unit as ADC, which is in $mm^2/s$. Their corresponding eigen vector is denoted as $\hat{\lambda_i}$ They are ordered from the greatest to the least diffusivity magnitude. $\hat{\lambda_1}$ points to the orientation of the greatest diffusion direction, and therefore $\lambda_1$ is also referred to as "longitudinal" or "axial" diffusivity. $\hat{\lambda_2}$ and $\hat{\lambda_3}$ are the lesser orthogonal diffusion directions, and their mean magnitude gives rise to the cross-sectional, or "radial" diffusivity.

\begin{figure}[ht]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=2.5in]{tensor.pdf}
\end{tabular}
\caption{Illustration of the eigen vectors of an ellipsoid tensor.}
\label{fig:tensor}
\end{center}
\end{figure}

@@

FA has been found to correlate with a wide spectrum of neuromorphological changes, including but not limited to multiple sclerosis \cite{Cercignani2001,Ciccarelli2003d,Miron2012}, vestibular schwannomas \cite{Chen2011b,Taoka2006,Wai2009}, tumors \cite{Mori2002,Schonberg2006,Nimsky2005}, schizophrenia \cite{Rametti2009,Fitzsimmons2009}, traumatic brain injury \cite{Bendlin2008,Kinnunen2011c,Hellyer2012}. RD is found to positively correlate with degree of demyelination in multiple sclerosis \cite{Song2005,Song2002,Klawiter2011,Janve2013}, and well as stages of white-matter development and aging \cite{Counsell2006,Sala2010}, including alzheimer's disease \cite{Acosta-Cabronero2012}. AD is correlated positively with axonal integrity \cite{Song2003}, and is found to be a correlate of axonal injury \cite{Budde2009} such as those found in traumatic brain injury \cite{Kinnunen2011c} and Gulf War illness \cite{Rayhan2013}. MD is correlated with tissue inflammation and edema \cite{Senda2012g}, and is found to correlate with diffusivity changes as a result of gliomas \cite{Stadlbauer2007}, multiple sclerosis \cite{Cercignani2001,Senda2012g,DeGroot2013}, amyotrophic lateral sclerosis \cite{Sharma2012}, and aging \cite{Kantarci2011}.

@@

In order to characterize the additional diffusion components in voxels with multiple fibres, changes in DWI acquisition strategies is necessary. A natural extension is to acquire additional diffusion directions. This type of acquisition strategies, which enable higher order diffusion modelling algorithms, are generally referred to as high angular resolution diffusion imaging (HARDI). The most wide-spread strategy is to equally distribute gradient directions as vertices of a triangle-tessellated geodesic sphere \cite{Frank2001}. Alternatively, it is also possible to distribute gradient directions in a rectilinear grid, which forms the basis of diffusion spectral imaging (DSI) \cite{Wedeen2005}. Additional acquisition strategies such as distributing gradient directions across multiple concentric spherical shells are also actively explored \cite{Jeurissen2014a}.

@@

The determination of gradient sampling pattern and B-value directly affect acquisition time and image quality. More complex gradient sampling and higher B-values can improve tractography quality. However the acquisition time can become prohibitively long in a clinical setting. Other factors that must be considered include the nature of patient pathology, motion artifacts, and the inclusion of other imaging modalities during clinical scan sessions.

@@

An end-to-end spherical deconvolution diffusion model was first proposed by Tournier et al. \cite{Tournier2004}. 
Consider a fiber population with distinct fiber components. The measured diffusion signal $S(\theta,\phi)$ can be written as:

\begin{equation}
S(\theta, \phi) = \sum_{i}^{} f_i \hat{A}_i R(\theta)
\end{equation}
where $\theta$ is the elevation angle,  $\phi$ the azimuth angle in spherical coordinates; $f_i$ is the volume fraction of the $i$-th fiber component, $\hat{A}$ is the rotation operator that orients fiber, and $R(\theta)$ is an axially symmetric response function.

@@

\begin{figure}[ht]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=100mm]{harmonics.pdf}
\end{tabular}
\caption{Visualization of the spherical harmonic functions up to l=3.}
\label{fig:harmonics}
\end{center}
\end{figure}

Spherical decovolution is considered to be the action of a set of rotations on a function over a sphere. Given $F^l $ as a vector of $l$th order spherical harmonic decomposition of  $F(\theta,\phi)$, with vector length $(2l + 1)$, and $R^l$ as the $l$th order rotational harmonic decomposite of $R(\theta)$, with vector length $(2l+1)^2$. The spherical harmonic representation of $S(\theta,\phi)$ is then:

\begin{equation}
S^l = R^l F^l
\end{equation}

To reduce the number of parameters to be computed, we can make the further assumption that the diffusion process is symmetric, and therefore the odd $l$ harmonic orders can be ignored, with only the even harmonics considered. 
The response function $R(\theta)$ can be directly estimated from the dataset by supplying DWI region with high anisotropy from a single fiber bundle, such as the body of the corpus callosum.

The maximum harmonic order $l_max$ that can be estimated is dependent on the obtained number of directions. For example, $l_max$ = 8 requires 45 directions, $l_max$ = 10 requires 66 directions, and so on. Larger $l_max$ are more sensitive to noise, which is a weakness for the first spherical deconvolution algorithm \cite{Tournier2004}.

\subsubsection{Constrained Spherical Deconvolution}
Further advances to the estimation method of the spherical deconvolution model made it more robust to noise \cite{Tournier2007b}. The noise of SD estimation is represented as spurious negative lobes in FOD, which are physically impossible, and do not represent any real phenomenon. To mitigate this, a non-negativity constraint is added. This method is referred to as \textit{constrained spherical deconvolution} (CSD).

The CSD constraint is applied iteratively using Tikhonov regularization \cite{Hansen1998}. The procedure involves obtaining an initial FOD estimate, and a set of directional peaks are then identified. The Tikhonov regularization then constraints the negative peaks to zero, and a new set of FOD and peaks are obtained. The process is applied iteratively until convergence, and the final constrained FOD is obtained.

The Tikhonov regularization minimizes the weighted sum with two terms:
\begin{equation}
\|{Af - b}^2\|^2 + \lambda^2 \| l(f - f')\|^2
\end{equation}
Where the first term is a linear least-square fit of solution $f$ to to data $b$ with problem matrix $A$. 
In terms of CSD, $f$ corresponds to the SH coefficients of the FOD, $b$ is the measured signal and $A$ is:
\begin{equation}
A = QR
\end{equation}
Where $Q$ is the SH coefficient to DW intensity mapping, and $R$ is the spherical deconvolution. 
The regularization term is the constrained matrix $L$ multiplying the norm of $f$ relative to an initial estimate $f'$. The parameter $\lambda$ controls the regularization amount, with $\lambda=0$ having no regularization, to $\lambda\rightarrow \infty $ making the solution tend to $f'$.

@@

CSD is robust to noise at relatively low B-values, and is capable of good quality FOD estimates at $B=1000 s/mm^2$. This means that existing DTI protocols can be extended for high order HARDI tractography by simply increasing the number of directions. With CSD, a 60 direction DWI acquisition can be used to confidently derive FOD with $l_{max}=8 $, and up to $l_{max}=12$ with super resolution enabled \cite{Tournier2007b} (Figure \ref{fig:csd}).

\begin{figure}[ht]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{csd-angles}
\end{tabular}
\caption{ Comparison of angle discrimination ability and CSD FOD. B-value=3000 \protect $s/mm^2$. Taken from \protect \cite{Tournier2007b} } 
\label{fig:csd}
\end{center}
\end{figure}

\subsection{Tractography}
Tractography is defined as the construction, or propagation of streamline from an underlying field of orientation information, in order to visualize continuous structures within a volume.  The orientation field is generally provided as a 3D matrix of directional information obtained from DTI, CSD, or other methods. For example, in DTI the directions are the eigen vector with the greatest magnitude $\hat{\lambda_1}$; for CSD, the directional vectors would be peaks from the FOD.

Applying pathway tracking to DWI imaging was first explored by Mori \cite{Mori2002b} and Basser\cite{Basser2000}. The family of methods that propagates by deterministic algorithms based on local orientation data is collectively called deterministic tractography. The first tracking algorithms were applied to DTI using step-wise polyline propagators that followed local DT tensors \cite{Conturo1999}. Further refinement of the line propagator to include neighbouring DT tensors, and interpolation of the local DT directions in order to obtain a smooth curve \cite{Mori2002b}. Basser introduced Runge-Kutta integration to allow more robust path-tracing when faced with noise in the local directions \cite{Basser2000}. Another similar path-integral method was also proposed by Tuch \cite{Tuch2000d}. The integrated tractography approaches are now standard practice for deterministic tractography \cite{Tournier2004,Pieper2004}.

Deterministic tractography lends itself naturally to anatomical interpretation and further geometric analysis. It was traditionally paired with DTI, and so was thought to be limited in the case of crossing fibers. Other methods attempted to circumvent the limitation with probabilistic approaches that include white matter tissue priors in the fiber propagation model. The most popular of the probabilistic approach is the FSL software's probabilistic tractography module (FDT) \cite{Behrens2007}, which also includes a model-based ball-and-stick diffusion model that can resolve 2-fiber crossings. The FDT software produces visitation images of its probabilistic streams that collate the probability of tract propagation at each voxel.  The visitation image, however, is hard to interpret and localize, and therefore does not rival deterministic methods in surgical visualization.

@@

\begin{figure}[ht]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{tractography1}
\end{tabular}
\caption{Example of cranial nerve reconstruction with deterministic tractography.} 
\label{fig:tract1}
\end{center}
\end{figure}

### 2017-12-02 thesis/contents/reviews/machine-learning.tex
checkpoint on edit

\section{Machine Learning}
Machine learning (ML) is a branch of computer science that aims to learn and form predictions from data. In contrast with static algorithmic programming, where the structure of the algorithms are predefined, machine learning approaches construct models based on observed data, in order to extend its predictions to the unobserved. Strategies and algorithms for constructing various models overlap with the fields of artificial intelligence, computational statistics, pattern recognition, and neuroscience. Advanced in modern machine learning, especially in deep neural networks \cite{LeCun2015}, in recent years have lead to wide spread adaptation of machine learning based online services in translation \cite{bahdanau2014neural}, image search \cite{simonyan2014very}, and speech recognition \cite{graves2013speech}. There is also surging interest in deploy machine learning systems in radiology and disease diagnosis in medicine.

Machine learning training strategies can be categorized into Supervised Learning, Unsupervised Learning, and Reinforcement Learning. Supervised learning involves training data that are pre-labeled. For example the data supplied can be a list of patients' age and gender, and the label can be whether they or not they are in pain. Unsupervised learning involve data without any prior labels, and the algorithm attempts to discover patterns within the data itself, and is often used in data mining. Reinforcement learning involves continuously learning with a reward function. It is generally used to train agents that have a continuous stream of data and needs to interact with an external environment, and is widely used in robotics.

\subsection{Supervised Learning}
Supervised learning is the mostly widely used method of training strategy. Most of the existing machine learning system in production are based primarily on supervised training with human-supplied data labels. Different algorithms can be used to fit a function that forms a decision boundary based on data features and labels. Possible models ranges from Decision Trees, linear models such as Support Vector Machines (SVM), Deep Neural Networks (DNN), and bayesian-based models such as Naive Bayes and Gaussian Process (Figure \ref{fig:ml-bounds}).

@@

\subsubsection{Gaussian Process}
The breakthrough results of deep convolutional neural networks in computer vision \cite{Krizhevsky2012} has sparked the interest of deploying similar systems for medical imaging diagnosis \cite{Greenspan2016}. However one disadvantage of DNN is in its inability to quantify uncertainty and provide interpretable models. Model interpretability is a fundamental requirement for machine learning in medicine, and therefore there is renewed interest in bayesian-based models like Gaussian Process \cite{gal2016dropout}. In fact it has been found that DNN with infinite number of hidden neurons approximates GP \cite{neal1996priors}.

Gaussian Process is defined \cite{rasmussen2006gaussian}) as a collection of random variables, any finite number of which have a joint Gaussian distribution. 
It is a distribution that is defined by a mean function $m(x)$ and covariance function $k(x,x') $ where
\begin{equation}
	\begin{split}
		m(x) &= \mathbb{E}[f(x)], \\
		k(x,x') &= \mathbb{E}[(f(x)-m(x)(f(x')-m(x'))],
	\end{split}
\end{equation}
And the Gaussian Process is defined as
\begin{equation}
f(x) \sim GP(m(x), k(x, x')) 
\end{equation}

@@

\subsubsection{Training}
The process of fitting a ML model on labeled data to form predictions is called training. The objective of the training process is to obtain a model that can both correctly predict observed labels, and is able to generalize its predictions to new data that are not part of the training set. Depending on the model algorithm, there can be many training parameter that govern the model outcome. The tuning of these parameters is referred to as hyper-parameter tuning. It is possible for certain hyper-parameters that result in model over-fitting, which is when a model is over-specialized in achieving the most accurate training label prediction, but perform poorly on unseen data. To properly evaluate model performance, a dataset is split into training/testing and validation sets, usually in 4:1 or 9:1 ratios. Where model training is performed solely on the training set, and evaluation of model performance is on the validation set to assess the ability for the model to generalize.

There are cases when the number of available data is low, and each data point may have a substantial impact on model performance. In these cases cross-validation is used. The idea is to divide the data into all possible combinations of training and validation splits, usually in the form of 5 fold to 10 fold. Then separate models are trained and evaluated on each combination, and their performance is averaged to form the final evaluation. This gives a reasonable approximate on the performance of the model when trained using all available data. Leave-one-out cross-validation is an extreme form of this: given a dataset of N items, each split will consist of 1 and N-1 items, which result in N possible combinations. It is used when the available data is extremely low.

@@

There has been a long history of applying statistical methods to the field of neuroimaging, from the very onset of the field. Neuroimaging studies often attempt to correlate physiological measures from neuro-activities with specific experimental variables or conditions. Neuro-activities may be measured from EGG sensors, or as voxels in functional or structural MR images. Traditionally data from each measured component was assumed to be independent. Correlative analysis were performed using general linear models, with post-hoc false-positive corrections. Therefore univariate models cannot take into account spatial or temporally correlative dynamics.

Increasing complexity in neuroimaging data have lead to widespread applications of multi-variable machine learning methods in neuroimaging. Application of ML in pain neuroimaging became controversial when Tor Wager \cite{Wager2013} demonstrated a Lasso-PCA model that classifies nociceptive pain with sensitivity and specificity of 90\% solely from fMRI imaging data. Additional studies showed consistently good performance of ML-based nociceptive pain discrimination. Discrimination of chronic pain from nociceptive pain proved to be more challenging, due to the high degree of similarity in pain-matrix activation patterns. The improvement in chronic pain detection lead to debates of the validity of presenting such methods as evidences in courtroom, where Davis et al argued against such actions \cite{Davis2012a}. While it is commonly accepted that pain activity in the brain involves multiple regions of the pain-matrix, evidences in some studies suggested that the posterior-insula is heavily involved in chronic pain. Segerdale et al claimed that the posterior insula is a primary pain region \cite{Segerdahl2015a}, thereby ignited renewed debates surrounding regional specific brain centers for chronic pain \cite{Davis2015}.

>>>>

\section{Machine Learning}
Machine learning (ML) is a branch of computer science that aims to learn and form predictions from data. In contrast with static algorithmic programming, where the structures of the algorithms are predefined, machine learning approaches construct models based on observed data, in order to extend its predictions to the unobserved. Strategies and algorithms for constructing various models overlap with the fields of artificial intelligence, computational statistics, pattern recognition, and neuroscience. Advanced in modern machine learning, especially in deep neural networks \cite{LeCun2015}, in recent years have led to wide spread adaptation of machine learning based online services in translation \cite{bahdanau2014neural}, image search \cite{simonyan2014very}, and speech recognition \cite{graves2013speech}. There is also surging interest in deploy machine learning systems in radiology and disease diagnosis in medicine\cite{Kamnitsas2016,Suk2014}.

Machine learning training strategies can be categorized into Supervised Learning, Unsupervised Learning, and Reinforcement Learning. Supervised learning involves training data that are pre-labelled. For example, the data supplied can be a list of patients' age and gender, and the label can be whether they or not they are in pain. Unsupervised learning involves data without any prior labels, and the algorithm attempts to discover patterns embedded within, and is often used in data mining. Reinforcement learning involves continuously learning with a reward function. It is generally used to train agents that have a continuous stream of data and needs to interact with an external environment and is widely used in robotics.

\subsection{Supervised Learning}
Supervised learning is the most widely used method of training strategy. Most of the existing machine learning systems in production are based primarily on supervised training with human-supplied data labels. Different algorithms can be used to fit a function that forms a decision boundary based on data features and labels. Possible models range from Decision Trees, linear models such as Support Vector Machines (SVM), Deep Neural Networks (DNN), and bayesian-based models such as Naive Bayes and Gaussian Process (Figure \ref{fig:ml-bounds}).

@@

\subsubsection{Gaussian Process}
The breakthrough results of deep convolutional neural networks in computer vision \cite{Krizhevsky2012} have sparked the interest of deploying similar systems for medical imaging diagnosis \cite{Greenspan2016}. However, one disadvantage of DNN is in its inability to quantify uncertainty and provide interpretable models. Model interpretability is a fundamental requirement for machine learning in medicine, and therefore there is renewed interest in bayesian-based models like Gaussian Process \cite{gal2016dropout}. DNN with an infinite number of hidden unit approximates GP \cite{neal1996priors}.

Gaussian Process is defined as a collection of random variables, any finite number of which have a joint Gaussian distribution \cite{rasmussen2006gaussian}. 
It is a distribution that is defined by a mean function $m(x)$ and covariance function $k(x,x') $ where
\begin{equation}
	\begin{split}
		m(x) &= \mathbb{E}[f(x)], \\
		k(x,x') &= \mathbb{E}[(f(x)-m(x)(f(x')-m(x'))],
	\end{split}
\end{equation}
And the Gaussian Process is defined as
\begin{equation}
f(x) \sim GP(m(x), k(x, x')) 
\end{equation}

@@

\subsubsection{Training}
The process of fitting an ML model on labelled data to form predictions is called training. The objective of the training process is to obtain a model that can both correctly predict observed labels, and can generalize its predictions to new data that are not part of the training set. Depending on the model algorithm, there can be many training parameters that govern the model outcome. The tuning of these parameters is referred to as hyper-parameter tuning. It is possible for models trained with certain hyper-parameter combinations to result in model over-fitting, which is when a model is over-specialized in achieving the most accurate training label prediction, but perform poorly on unseen data. To properly evaluate model performance, a dataset is split into training/testing and validation sets, usually in 4:1 or 9:1 ratios, where model training is performed solely on the training set, and evaluation of model performance is on the validation set to assess the ability of the model to generalize.

There are instances when the number of available data is low, and each data point may have a substantial impact on model performance. In these cases, cross-validation is applied. The purpose is to divide the data into all possible combinations of training and validation splits, usually in the form of 5 fold to 10 fold. Then separate models are trained and evaluated on each combination, and their performance is averaged to form the final evaluation. This gives a reasonable approximate on the performance of the model when trained using all available data. Leave-one-out cross-validation is a special form where only one item is used for the validation split: given a dataset of N items, each split will consist of 1 and N-1 items, which result in N possible combinations. It is used when the available data are few.

@@

There is an extensive history of applying statistical methods to the field of neuroimaging, from the very onset of the field. Neuroimaging studies often attempt to correlate physiological and image-based measures with neuro-activities with specific experimental variables or conditions. Neuro-activities may be measured from EGG sensors, or as voxels in functional or structural MR images. Traditionally data from each measured feature component was assumed to be independent. These components maybe individual sensor readings, or voxel-based measures. Correlative analysis were performed using general linear models, with post-hoc false-positive corrections. This type of univariate analysis do not have the ability to discover key relationships between feature components. Therefore univariate models cannot take into account spatial or temporally correlative dynamics.

Increasing complexity in neuroimaging data has lead to widespread applications of multi-variable machine learning methods in neuroimaging. Application of ML in pain neuroimaging became controversial when Tor Wager \cite{Wager2013} demonstrated a Lasso-PCA model that classifies nociceptive pain with sensitivity and specificity of 90\% solely from fMRI imaging data. Additional studies showed consistently good performance of ML-based nociceptive pain discrimination. Discrimination of chronic pain from nociceptive pain proved to be more challenging, due to the high degree of similarity in pain-matrix activation patterns. The improvement in chronic pain detection lead to debates of the validity of presenting such methods as evidences in courtroom, where Davis et al argued against such actions \cite{Davis2012a}. While it is commonly accepted that pain activity in the brain involves multiple regions of the pain-matrix, evidences in some studies suggested that the posterior-insula is heavily involved in chronic pain. Segerdale et al claimed that the posterior insula is a primary pain region \cite{Segerdahl2015a}, thereby ignited renewed debates surrounding regional specific brain centers for chronic pain \cite{Davis2015}.
